Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2017-06-08 Master Thesis/synopsis.tex
Fix typo

Let $w_1,w_2 \in\operatorname{Spv}_\text{d}^{(2)}(K)$ be two different valuations. Then the following statements are equivalent:
\begin{enumerate}[(i)]
\item The composite $K_{w_1}K_{w_2}$ is not algebraically closed.
\item $w_1,w_2\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ and $v(w_1) = v(w_2)$.
\end{enumerate}
The implication (i)$\implies$(ii) follows from Lemma \ref{2.2}. Now assume (ii) holds. Then $k(v)$ is separably closed where $v=v(w_1)$. By Lemma \ref{lemma-henselization-specialization} we have $K_{w_1}K_{w_2}\subset K_v^\text{t}$ and by Theorem \ref{theorem-absoluteinertia} (ii) $K_{w_1}K_{w_2}=K_v^\text{t}$. The extension of $v$ on $K_v^\text{t}$ has value group $\Gamma_v$, which is finitely generated. However, value groups of valuations on an algebraically closed field are divisible, hence $K_v^\text{t}$ can't be algebraically closed.

Furthermore, in this case, $K'_{w'_1}K'_{w'_2} = (K_{w_1}K_{w_2})' \neq \overline{K'}$ and again by Lemma \ref{2.2} we get: \[w'_1,w'_2\in\operatorname{Spv}_\text{rd}^{(2)}(K)\quad \text{and} \quad v(w'_1) = v(w'_2)\]
Thus $\sigma: \operatorname{Spv}_\text{d}^{(2)}(K)\to \operatorname{Spv}_\text{d}^{(2)}(K')$ maps $\operatorname{Spv}_\text{rd}^{(2)}(K)$ onto $\operatorname{Spv}_\text{rd}^{(2)}(K')$ such that:
\[ v(w_1)=v(w_2) \iff v(w'_1) = v(w'_2) \]
Let $v\in\operatorname{Spv}_\text{d}^{(1)}(K)$. Since $v\not\in |\operatorname{Spv}(K)|$, there exists a $w\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ such that $v(w)=v$. We define $v'=v(w')$. This shows that $\sigma$ also induces a bijection: 
\[\sigma:\operatorname{Spv}_\text{d}^{(1)}(K)\to\operatorname{Spv}_\text{d}^{(1)}(K'),\ v\mapsto v'\] 
This map is well-defined and for all $w\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ we have $\sigma(v(w)) = v(\sigma(w))$.

Let $v\in\operatorname{Spv}_\text{d}^{(1)}(K)$ and $w\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ with $v(w)=v$. Then $k(v)$ is finitely generated of dimension $1$ by Lemma \ref{rank-dimension-inequality}, thus a global field. By Theorem \ref{theorem-absoluteinertia}, we have: 
\[G_{K_w}/N \cong \operatorname{Gal}(K_v^\text{t}/K_w)\cong G_{k(v_w)}\] 
where $v_w$ is the unique extension of $v$ on $K_w$ and $N=G_{K_v^\text{t}}$. With Lemma \ref{2.2-pre} we can see that $k(v_w)$ is the henselization of $k(v)$ with respect to $w\mod v\in\operatorname{Spv}k(v)$. Let $H$ be the closed subgroup of $G_{K_v}$ generated by all subgroups $G_{K_w}$ for $w\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ with $v(w) = v$. Since a global field has no non-trivial algebraic extension in which every prime splits completely, we get $H/N= G_{K_v}/H\cong G_{k(v)}$, thus $H=G_{K_v}$. Since $(K_w)'=K'_{w'}$, this shows that $(K_v)'$ is a henselization of $K'$ with respect to $v'=\sigma(v)$. We write $K'_{v'} = (K_v)'$.

Furthermore, let $w_1,w_2\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ with $w_1\neq w_2$ and $v=v(w_1)=v(w_2)$, then:
\[ \sigma(G_{K^\text{t}_v}) = \sigma(G_{K_{w_1}}\cap G_{K_{w_2}}) = G_{K'_{w'_1}}\cap G_{K'_{w'_2}} = G_{(K')^\text{t}_{v'}} \]
Therefore, $\sigma$ induces with Theorem \ref{theorem-absoluteinertia} the following isomorphism:
\[ G_{k(v)} \cong G_{K_v}/G_{K_v^\text{t}} \to G_{K'_{v'}}/G_{(K')_{v'}^\text{t}}\cong G_{k(v')} \]
The rest follows from the Theorem of Neukirch-Uchida \ref{thm:neukirch-uchida}.
\end{proof}

>>>>

Let $w_1,w_2 \in\operatorname{Spv}_\text{d}^{(2)}(K)$ be two different valuations. Then the following statements are equivalent:
\begin{enumerate}[(i)]
\item The composite $K_{w_1}K_{w_2}$ is not algebraically closed.
\item $w_1,w_2\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ and $v(w_1) = v(w_2)$.
\end{enumerate}
The implication (i)$\implies$(ii) follows from Lemma \ref{2.2}. Now assume (ii) holds. Then $k(v)$ is separably closed where $v=v(w_1)$. By Lemma \ref{lemma-henselization-specialization} we have $K_{w_1}K_{w_2}\subset K_v^\text{t}$ and by Theorem \ref{theorem-absoluteinertia} $K_{w_1}K_{w_2}=K_v^\text{t}$. The extension of $v$ on $K_v^\text{t}$ has value group $\Gamma_v$, which is finitely generated. However, value groups of valuations on an algebraically closed field are divisible, hence $K_v^\text{t}$ can't be algebraically closed.

Furthermore, in case (ii), $K'_{w'_1}K'_{w'_2} = (K_{w_1}K_{w_2})' \neq \overline{K'}$ follows from (i) and again by Lemma \ref{2.2} we get: \[w'_1,w'_2\in\operatorname{Spv}_\text{rd}^{(2)}(K)\quad \text{and} \quad v(w'_1) = v(w'_2)\]
Thus $\sigma: \operatorname{Spv}_\text{d}^{(2)}(K)\to \operatorname{Spv}_\text{d}^{(2)}(K')$ maps $\operatorname{Spv}_\text{rd}^{(2)}(K)$ onto $\operatorname{Spv}_\text{rd}^{(2)}(K')$ such that:
\[ v(w_1)=v(w_2) \iff v(w'_1) = v(w'_2) \]
Let $v\in\operatorname{Spv}_\text{d}^{(1)}(K)$. Since $v\not\in |\operatorname{Spv}(K)|$, there exists a $w\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ such that $v(w)=v$. We define $v'=v(w')$. This shows that $\sigma$ also induces a bijection: 
\[\sigma:\operatorname{Spv}_\text{d}^{(1)}(K)\to\operatorname{Spv}_\text{d}^{(1)}(K'),\ v\mapsto v'\] 
This map is well-defined and for all $w\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ we have $\sigma(v(w)) = v(\sigma(w))$.

Let $v\in\operatorname{Spv}_\text{d}^{(1)}(K)$ and $w\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ with $v(w)=v$. Then $k(v)$ is finitely generated of dimension $1$ by Lemma \ref{rank-dimension-inequality}, thus a global field. By Theorem \ref{theorem-absoluteinertia}, we have: 
\[G_{K_w}/N \cong \operatorname{Gal}(K_v^\text{t}/K_w)\cong G_{k(v_w)}\] 
where $v_w$ is the unique extension of $v$ on $K_w$ and $N=G_{K_v^\text{t}}$. With Lemma \ref{2.2-pre} we can see that $k(v_w)$ is the henselization of $k(v)$ with respect to $w\mod v\in\operatorname{Spv}k(v)$. Let $H$ be the closed subgroup of $G_{K_v}$ generated by all subgroups $G_{K_w}$ for $w\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ with $v(w) = v$. Since a global field has no non-trivial algebraic extension in which every prime splits completely, we get $H/N= G_{K_v}/N\cong G_{k(v)}$, thus $H=G_{K_v}$. Since $(K_w)'=K'_{w'}$, this shows that $(K_v)'$ is a henselization of $K'$ with respect to $v'=\sigma(v)$. We write $K'_{v'} = (K_v)'$.

Now let $w_1,w_2\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ with $w_1\neq w_2$ and $v=v(w_1)=v(w_2)$. We have:
\[ \sigma(G_{K^\text{t}_v}) = \sigma(G_{K_{w_1}}\cap G_{K_{w_2}}) = G_{K'_{w'_1}}\cap G_{K'_{w'_2}} = G_{(K')^\text{t}_{v'}} \]
Therefore, $\sigma$ induces with Theorem \ref{theorem-absoluteinertia} the following isomorphism:
\[ G_{k(v)} \cong G_{K_v}/G_{K_v^\text{t}} \to G_{K'_{v'}}/G_{(K')_{v'}^\text{t}}\cong G_{k(v')} \]
The rest follows from the Theorem of Neukirch-Uchida \ref{thm:neukirch-uchida}.
\end{proof}
